---
title: "Income"
author: "Darius Kharazi"
date: "9/10/2017"
output: pdf_document
---

## Import Statements

```{r import, include=FALSE}
## Import Libraries 
library(knitr)
library(readr)
library(ggplot2)

## Import Data
income <- read_csv("~/Downloads/income/income_tr.csv") # Input location of "income_tr.csv" file
income.df <- data.frame(income)

## Create new data frame
newincome.df <- data.frame(income.df$ID)
newincome.df[,2:42] <- 0

## Rename columns
colnames(newincome.df) <- c("ID", "isAge10-19", "isAge20-29", "isAge30-39", "isAge40-49", "isAge50-59", "isAge60-69", "isAge70-79", "isAge80-89", "isWorkPriv", "isWorkOther", "isWgt20-99k", "isWgt100-199k", "isWgt200-299k", "isWgt300-399k", "isWgt400-499k", "isWgt500k+", "isEduHighGradOrLess", "isEduSCol-Bach", "isEduMastOrMore", "isMarried", "isNeverMarried", "isDivOrSep", "isWidowed", "isOccBus", "isOccTech", "isOccOther", "isRelHusbOrWife", "isRelUnmarried", "isRelOther", "isRaceWhite", "isRaceOther", "isMale", "isFemale", "isCapGain0", "isCapGainOther", "isCapLoss0", "isCapLossOther", "isHPW40OrLess", "isHPW41OrMore", "isCountryUS", "isCountryOther")
```

## Exploratory Analysis

```{r margen}
## Analyze Class
nrow(subset(income.df, income.df$class==">50K"))
nrow(subset(income.df, income.df$class=="<=50K"))

## Analyze Gender
barplot(table(income.df$gender))

## Analyze Education and Marriage
somecollege <- subset(newincome.df, newincome.df$isEduHighGradOrLess == 1)
bachelors <- subset(newincome.df, newincome.df$`isEduSCol-Bach` == 1)
masters <- subset(newincome.df, newincome.df$isEduMastOrMore == 1)

sum(somecollege$isMarried)/nrow(somecollege)
sum(bachelors$isMarried)/nrow(bachelors)
sum(masters$isMarried)/nrow(masters)

nrow(somecollege)
nrow(bachelors)
nrow(masters)

boxplot(income.df$education_cat~income.df$class)
```

After analysis of the makeup of the dataset’s native country variable, it seems that 90% of individuals are from the United States, which is extremely high and may indicate reason for variable manipulation. Also, nearly 90% of individuals indicated that they had a 0% capital gain, and nearly 95% of individuals indicated that they had a 0% capital loss. These large quantities may also indicate reason for variable manipulation. Lastly, it seems that there is a large division amongst individuals who work forty hours per week (or less) and more than 40 hours per week. There also seems to be a few outliers for people who work close to 100 hours per week. For both of these reasons, it may be reasonable to perform analysis differentiating the two groups with some sort of variable transformation.

The dataset seems to include interesting conditional information, as well, which may be important to consider during further analysis. For example, the dataset contains nearly double the amount of men compared to women. Also, there is a smaller percentage of people married that have a college education, but less than a master’s degree. However, there is a higher percentage of people married who have never been to college. The highest percentage of people married are those who have a master's degree or greater, but that could be a result of a small sample size. Also, nearly 43% of people who are white are married in the dataset; whereas nearly 30% of people who aren't white are married in the dataset. This could also be a result of the smaller sample size of people who are not white. 

```{r}
## Analyze Race and Marriage
whiterace <- subset(newincome.df, newincome.df$isRaceWhite == 1)
otherrace <- subset(newincome.df, newincome.df$isRaceOther == 1)

sum(whiterace$isMarried)/nrow(whiterace)
sum(otherrace$isMarried)/nrow(otherrace)

nrow(whiterace)
nrow(otherrace)

## Analyze Race and Occupation
whiterace <- subset(newincome.df, newincome.df$isRaceWhite == 1)
otherrace <- subset(newincome.df, newincome.df$isRaceOther == 1)

sum(whiterace$isOccBus)/nrow(whiterace)
sum(otherrace$isOccBus)/nrow(otherrace)

nrow(whiterace)
nrow(otherrace)

## Analyze Native Country and Education
UScountry <- subset(newincome.df, newincome.df$isCountryUS == 1)
othercountry <- subset(newincome.df, newincome.df$isCountryOther == 1)

sum(UScountry$isEduHighGradOrLess)/nrow(UScountry)
sum(othercountry$isEduHighGradOrLess)/nrow(othercountry)
sum(UScountry$`isEduSCol-Bach`)/nrow(UScountry)
sum(othercountry$`isEduSCol-Bach`)/nrow(othercountry)
sum(UScountry$isEduMastOrMore)/nrow(UScountry)
sum(othercountry$isEduMastOrMore)/nrow(othercountry)

nrow(UScountry)
nrow(othercountry)

## Analyze Age
age10 <- subset(newincome.df, newincome.df$`isAge10-19` == 1)
age20 <- subset(newincome.df, newincome.df$`isAge20-29` == 1)
age30 <- subset(newincome.df, newincome.df$`isAge30-39` == 1)
age40 <- subset(newincome.df, newincome.df$`isAge40-49` == 1)
age50 <- subset(newincome.df, newincome.df$`isAge50-59` == 1)
age60 <- subset(newincome.df, newincome.df$`isAge60-69` == 1)
age70 <- subset(newincome.df, newincome.df$`isAge70-79` == 1)
age80 <- subset(newincome.df, newincome.df$`isAge80-89` == 1)

nrow(age10)
nrow(age20)
nrow(age30)
nrow(age40)
nrow(age50)
nrow(age60)
nrow(age70)
nrow(age80)

boxplot(income.df$age~income.df$class)
```

Additionally, nearly 37% of people who are white have a business-related occupation, and 31% of people who are not white have a business-related occupation. Again, the slight difference could be a result of the smaller sample size of people who are not white. Lastly, 45% of both people whose native country is the United States and people whose native country is not the United States have less than a college-level education. Furthermore, nearly 7% of people whose native country is the United States have a master’s degree or further education; whereas nearly 16% of people whose native country is not the United States have a master’s degree or further education. Therefore, nearly 47% of people from the United States have some sort of college education, but less education than a master’s degree; and only around 38% of people who are not from the United States have some sort of college education, but less education than a master’s degree.

```{r fiftyclass}
## Analyze Class
nrow(subset(income.df, income.df$class==">50K"))
nrow(subset(income.df, income.df$class=="<=50K"))
```

The "class" variable seemed to be a good indicator of income under some analysis and assumption. There needed to be a large enough sample size in each class, ">50K" and "<=50K," in order to ensure any analysis could be conducted. Although only 100 entries existed with class equal to ">50K," and 420 entries existed with class equal to "<=50K," we should assume that the sample sizes for each group are large enough in order to perform further analysis. Most likely, the skewness is a result of the actual data and not the collection of data, since fewer people are wealthier than the given class.

```{r classmar}
## Analyze Class and Marital Status
ggplot(income.df, aes(income.df$class, ..count..)) + 
  geom_bar(aes(fill = income.df$marital_status), position = "dodge")

table(income.df$marital_status)/nrow(income.df)
```

After analysis of the distribution of classes given an individual’s marital status, it seems that a smaller percentage of people from the “>50K” class have never been married; whereas the largest percentage of people from the “<=50K” class have never been married.

```{r classwork}
## Analyze Class and Workclass
ggplot(income.df, aes(income.df$class, ..count..)) + 
  geom_bar(aes(fill = income.df$workclass), position = "dodge")

table(income.df$workclass)/nrow(income.df)
```

This could indicate some potential interaction between the “marital status” variable and “class” variable. Additionally, the largest percentage of individuals from each class are part of the “private class,” which seems to not indicate any interaction occurring between the two classes in the scope of an individual’s work class.

```{r classocc}
## Analyze Class and Occupation
ggplot(income.df, aes(income.df$class, ..count..)) + geom_bar(aes(fill = income.df$occupation), position = "dodge")

table(income.df$workclass)/nrow(income.df)
```

However, it seems that nearly 70% of individuals indicated that they are "private class." Therefore, it seems reasonable to divide individuals by "private class" and "other class," rather than including the numerous amounts of work classes in the analysis. Also, it seems like a higher percentage of individuals from the ">50K" class are "Prof-specialty" and "Exec-managerial," compared to the "<=50K" class. The higher percentage of individuals from the "<=50K" class are "Adm-clerical" and "Other-
service".

```{r classrel}
## Analyze Class and Relationship
ggplot(income.df, aes(income.df$class, ..count..)) +
  geom_bar(aes(fill = income.df$relationship), position = "dodge")

table(income.df$relationship)/nrow(income.df)
```

This finding may be a consequence of the differing sample sizes between the classes, but could indicate an interaction between the variables. Furthermore, the highest percentage of people from class “>50K” consider their relationship status to be either “husband” or “wife;” whereas the highest percentage of people considered themselves “not-in-family” from the class “<=50K.” Similar to other variables, this may indicate an interaction between relationship and class.

```{r classrace}
## Analyze Class and Race
ggplot(income.df, aes(income.df$class, ..count..)) +
  geom_bar(aes(fill = income.df$race), position = "dodge")

table(income.df$race)/nrow(income.df)
```

Interestingly, the highest percentage of race from both classes is “white.” However, nearly 83% of individuals indicated that they are white; and, therefore, it seems reasonable to divide individuals into either “white race” or “other race” for our analysis.

## Preprocessing

After loading in the required libraries and data frames, we want to factor some of the variables into more appropriate categories.

```{r age, include=FALSE}
## Get indices of corresponding entries for "age"
tenInd <- which(income.df$age >= 10 & income.df$age <= 19)
twentyInd <- which(income.df$age >= 20 & income.df$age <= 29)
thirtyInd <- which(income.df$age >= 30 & income.df$age <= 39)
fortyInd <- which(income.df$age >= 40 & income.df$age <= 49)
fiftyInd <- which(income.df$age >= 50 & income.df$age <= 59)
sixtyInd <- which(income.df$age >= 60 & income.df$age <= 69)
seventyInd <- which(income.df$age >= 70 & income.df$age <= 79)
eightyInd <- which(income.df$age >= 80 & income.df$age <= 89)

## Assign correct values to newincome.df according to "age"
newincome.df$`isAge10-19`[tenInd] = 1
newincome.df$`isAge20-29`[twentyInd] = 1
newincome.df$`isAge30-39`[thirtyInd] = 1
newincome.df$`isAge40-49`[fortyInd] = 1
newincome.df$`isAge50-59`[fiftyInd] = 1
newincome.df$`isAge60-69`[sixtyInd] = 1
newincome.df$`isAge70-79`[seventyInd] = 1
newincome.df$`isAge80-89`[eightyInd] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isAge10-19 if the individual is between the ages of 10 and 19 (inclusive);  
- isAge20-29 if the individual is between the ages of 20 and 29 (inclusive);  
- isAge30-39 if the individual is between the ages of 30 and 39 (inclusive);  
- isAge40-49 if the individual is between the ages of 40 and 49 (inclusive);  
- isAge50-59 if the individual is between the ages of 50 and 59 (inclusive);  
- isAge60-69 if the individual is between the ages of 60 and 69 (inclusive);  
- isAge70-79 if the individual is between the ages of 70 and 79 (inclusive);  
- isAge80-89 if the individual is between the ages of 80 and 89 (inclusive);  
- Otherwise, the values will be 0.

The question is whether recoding missing values into valid 0 is warranted. For binary (0 vs 1) data 0 means "absent" while missing means "absent or present - not known". No similarity measure itself can help you decide how to go about missing data.

```{r workclass, include=FALSE}
## Get indices of corresponding entries for "workclass"
privInd <- which(income.df$workclass %in% c("Private"))
othInd <- which(!income.df$workclass %in% c("Private"))

## Assign correct values to newincome.df according to "workclass"
newincome.df$isWorkPriv[privInd] = 1
newincome.df$isWorkOther[othInd] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isWorkPriv if the individual's workclass is private;
- isWorkOther if the individual's workclass is not private.
- Otherwise, the values will be 0.

```{r wgt, include=FALSE}
## Get indices of corresponding entries for "fnlwgt"
twentyInd <- which(income.df$fnlwgt >= 20000 & income.df$fnlwgt < 100000)
onehundInd <- which(income.df$fnlwgt >= 100000 & income.df$fnlwgt < 200000)
twohundInd <- which(income.df$fnlwgt >= 200000 & income.df$fnlwgt < 300000)
threehundInd <- which(income.df$fnlwgt >= 300000 & income.df$fnlwgt < 400000)
fourhundInd <- which(income.df$fnlwgt >= 400000 & income.df$fnlwgt < 500000)
fivehundInd <- which(income.df$fnlwgt >= 500000 & income.df$fnlwgt < 1000000)

## Assign correct values to newincome.df according to "fnlwgt"
newincome.df$`isWgt20-99k`[twentyInd] = 1
newincome.df$`isWgt100-199k`[onehundInd] = 1
newincome.df$`isWgt200-299k`[twohundInd] = 1
newincome.df$`isWgt300-399k`[threehundInd] = 1
newincome.df$`isWgt400-499k`[fourhundInd] = 1
newincome.df$`isWgt500k+`[fivehundInd] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isWft20-99k if the individual has a "fnlwgt" between 20,000 and 100,000 (exclusive);
- isWft100-199k if the individual has a "fnlwgt" between 100,000 and 200,000 (exclusive);
- isWft200-299k if the individual has a "fnlwgt" between 200,000 and 300,000 (exclusive);
- isWft300-399k if the individual has a "fnlwgt" between 300,000 and 400,000 (exclusive);
- isWft400-499k if the individual has a "fnlwgt" between 400,000 and 500,000 (exclusive);
- isWft500k+ if the individual has a "fnlwgt" of 500,000 or greater.
- Otherwise, the values will be 0.

```{r education, include=FALSE}
## Get indices of corresponding entries for "education"
highInd <- which(income.df$education %in% c("Preschool", "1st-4th", "5th-6th", "7th-8th", 
                                            "9th", "10th", "11th", "12th", "HS-grad"))
bachInd <- which(income.df$education %in% c("Some-college", "Assoc-acdm", "Assoc-voc",
                                             "Bachelors"))
mastInd <- which(income.df$education %in% c("Masters", "Doctorate", "Prof-school"))

## Assign correct values to newincome.df according to "education"
newincome.df$isEduHighGradOrLess[highInd] = 1
newincome.df$`isEduSCol-Bach`[bachInd] = 1
newincome.df$isEduMastOrMore[mastInd] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isEduHighGradOrLess if the individual has a high school education or less;
- isEduSCol-Bach if the individual has education between some college and a bachelors degree;
- isEduMastOrMore if the individual has a Masters degree or more.
- Otherwise, the values will be 0.

Also, the variable "education_cat" was excluded from the newly created data frame, since the variable is essentially equivalent to the "education" variable in the "income.df" data frame.

```{r maritalstatus, include=FALSE}
## Get indices of corresponding entries for "marital_status"
marriedInd <- which(income.df$marital_status %in% c("Married-AF-spouse", 
                                                    "Married-civ-spouse", 
                                                    "Married-spouse-absent"))
nMarriedInd <- which(income.df$marital_status %in% c("Never-married"))
divInd <- which(income.df$marital_status %in% c("Divorced", "Separated"))
widInd <- which(income.df$marital_status %in% c("Widowed"))

## Assign correct values to newincome.df according to "marital_status"
newincome.df$isMarried[marriedInd] = 1
newincome.df$isNeverMarried[nMarriedInd] = 1
newincome.df$isDivOrSep[divInd] = 1
newincome.df$isWidowed[widInd] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isMarried if the individual is married;
- isNeverMarried if the individual has never been married;
- isDivOrSep if the individual is divorced or separated;
- isWidowed if the individual is a widow.
- Otherwise, the values will be 0.

```{r occupation, include=FALSE}
## Get indices of corresponding entries for "occupation"
occBusInd <- which(income.df$occupation %in% c("Adm-clerical", 
                                                    "Exec-managerial", 
                                                    "Sales"))
occTechInd <- which(income.df$occupation %in% c("Machine-op-inspct", "Tech-support"))
occOthInd <- which(income.df$occupation %in% c("?", "Craft-repair", "Farming-fishing",
                                               "Handlers-cleaners", "Other-service", 
                                               "Priv-house-serv", "Prof-specialty", 
                                               "Protective-serv", "Transport-moving"))

## Assign correct values to newincome.df according to "occupation"
newincome.df$isOccBus[occBusInd] = 1
newincome.df$isOccTech[occTechInd] = 1
newincome.df$isOccOther[occOthInd] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isOccBus if the individual has an occupation relating to business;
- isOccTech if the individual has an occupation relating to technology;
- isOccOther if the individual has an occupation that doesn't relate to business or technology.
- Otherwise, the values will be 0.


```{r relationship, include=FALSE}
## Get indices of corresponding entries for "relationship"
relHW <- which(income.df$relationship %in% c("Husband", 
                                                    "Wife"))
relUnmarr <- which(income.df$relationship %in% c("Unmarried"))
relOther <- which(income.df$relationship %in% c("Not-in-family", "Other-relative",
                                                "Own-child"))

## Assign correct values to newincome.df according to "relationship"
newincome.df$isRelHusbOrWife[relHW] = 1
newincome.df$isRelUnmarried[relUnmarr] = 1
newincome.df$isRelOther[relOther] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isRelHusbOrWife if the individual either considers themselves as a husband or wife;
- isRelUnmarried if the individual is unmarried;
- isRelOther if the individual doesn't consider themselves as married or unmarried.
- Otherwise, the values will be 0.

```{r race, include=FALSE}
## Get indices of corresponding entries for "race"
raceWhite <- which(income.df$race %in% c("White"))
raceOther <- which(!income.df$race %in% c("White"))

## Assign correct values to newincome.df according to "race"
newincome.df$isRaceWhite[raceWhite] = 1
newincome.df$isRaceOther[raceOther] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isRaceWhite if the individual is white;
- isRaceOther if the individual is not white.
- Otherwise, the values will be 0.

```{r gender, include=FALSE}
## Get indices of corresponding entries for "gender"
genMale <- which(income.df$gender %in% c("Male"))
genFemale <- which(income.df$gender %in% c("Female"))

## Assign correct values to newincome.df according to "gender"
newincome.df$isMale[genMale] = 1
newincome.df$isFemale[genFemale] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isMale if the individual is male;
- isFemale if the individual is female.
- Otherwise, the values will be 0.

```{r capgain, include=FALSE}
## Get indices of corresponding entries for "capital_gain"
gainZero <- which(income.df$capital_gain == 0)
gainOther <- which(income.df$capital_gain != 0)

## Assign correct values to newincome.df according to "capital_gain"
newincome.df$isCapGain0[gainZero] = 1
newincome.df$isCapGainOther[gainOther] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isCapGain0 if the individual has a capital gain that equals 0;
- isCapGainOther if the individual has a capital gain that is greater than 0.
- Otherwise, the values will be 0.

```{r caploss, include=FALSE}
## Get indices of corresponding entries for "capital_loss"
lossZero <- which(income.df$capital_loss == 0)
lossOther <- which(income.df$capital_loss != 0)

## Assign correct values to newincome.df according to "capital_loss"
newincome.df$isCapLoss0[lossZero] = 1
newincome.df$isCapLossOther[lossOther] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isCapLoss0 if the individual has a capital loss that equals 0;
- isCapLossOther if the individual has a capital loss that is greater than 0.
- Otherwise, the values will be 0.

```{r hrsperweek, include=FALSE}
## Get indices of corresponding entries for "hours_per_week"
hrsForty <- which(income.df$hour_per_week > 0 & income.df$hour_per_week <= 40)
hrsGreater <- which(income.df$hour_per_week > 40 & income.df$hour_per_week <= 100)

## Assign correct values to newincome.df according to "hours_per_week"
newincome.df$isHPW40OrLess[hrsForty] = 1
newincome.df$isHPW41OrMore[hrsGreater] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isHPW40OrLess if the individual works 40 hours per week or less;
- isHPW41OrMore if the individual works 41 hours per week or more.
- Otherwise, the values will be 0.

```{r country, include=FALSE}
## Get indices of corresponding entries for "native_country"
countryUS <- which(income.df$native_country %in% c("United-States"))
countryOther <- which(!income.df$native_country %in% c("United-States"))

## Assign correct values to newincome.df according to "native_country"
newincome.df$isCountryUS[countryUS] = 1
newincome.df$isCountryOther[countryOther] = 1
```

The newly created data frame will be assigned the value 1 to each of the following variables:

- isCountryUS if the individual's native country is the United States;
- isCountryOther if the individual's native country is anywhere that isn't the United States.
- Otherwise, the values will be 0.

## Description of the kNN Implementation

After exploratory analysis, data preprocessing needed to be performed. In the early stages of program implementation, it made sense to perform variable manipulation, primarily due to the reasons mentioned in the exploratory analysis. Since certain categorical variables contained such high percentages of the total makeup at certain levels, it seemed reasonable to transform these variables. Also, variable transformation seemed reasonable for variables that contained quite a few outliers, such as the “fnlwgt” and “hours per week” variables, perhaps to understand the outliers more and capture this information. Lastly, since so many data entries had missing values, data deletion was not much of an option. Manipulating and creating a new “other” level for certain variables, such as “occupation,” seemed like a reasonable approach for handling missing values. Because of the huge quantity of variables that needed their own variables for distinct levels, the variable manipulation process engendered binary data.

After data preprocessing, cosine and jaccard similarity matrices and cosine and jaccard output matrices are created for two separate purposes. The cosine and jaccard similarity matrices are two symmetrical matrices with the diagonal values equal to 1. Their purpose is to demonstrate the cosine and jaccard similarity values for each data entry in comparison with other data entries. After these similarity matrices are calculated, the output matrices poll the corresponding data and identify the k-most similar data entries in relation to every other data entry. The parameter “k” is adjustable, and determines the limit of indices and proximities that should be included in the output matrix. Each proximity that is included in the output matrix provides its corresponding index from the income dataset, as well.

Since each variable contained binary data after the variable transformation process, it seemed reasonable to calculate cosine and jaccard proximities for each data entry. Cosine similarity seemed reasonable for analysis of each data entry’s vector in the cosine similarity matrix. Additionally, calculating jaccard similarities seemed reasonable given our transformed binary data, since jaccard similarities should only be calculated on binary data.

## Cosine Similarity Function

```{r cossim, include=FALSE}
## Convert data frame to matrix
newincome.m <- data.matrix(newincome.df)

## Create adjustable k
k <- 5 # Adjust k here

## Create a new data frame containing cosine similarities and output values
cossim.m <- matrix(nrow=(nrow(newincome.df)+1), ncol=(nrow(newincome.df)+1))
cossim.m[2:nrow(cossim.m),1] <- newincome.df$ID
cossim.m[1,2:nrow(cossim.m)] <- newincome.df$ID
cosine.df <- data.frame(income.df$ID)
cosine.df[,2:((k*2)+1)] <- 0

## Rename columns
names(cosine.df)[seq(2,ncol(cosine.df),2)] <- "Index"
names(cosine.df)[seq(1,ncol(cosine.df),2)] <- "Proximity"
names(cosine.df)[1] <- "ID"

## Create a function for calculating cosine similarity of two distinct data entries
cosine <- function(indi, indj) {
  xVec <- as.numeric(newincome.df[indi,2:42])
  yVec <- as.numeric(newincome.df[indj,2:42])
  xlen <- sqrt(sum(xVec^2))
  ylen <- sqrt(sum(yVec^2))
  simC <- sum(t(xVec)*yVec)/(xlen*ylen)
  return(simC)
}

## Perform function for each entry
for (rowi in 1:nrow(newincome.df)) {
  for (rowj in 1:nrow(newincome.df)) {
    cossim.m[rowi+1, rowj+1] <- cosine(rowi, rowj)
  }
}

## Convert matrix to dataframe
cossim.df <- data.frame(cossim.m)
cossim.df[1,1] = 0
colnames(cossim.df) <- cossim.df[1,]
cossim.df <- cossim.df[-1,]
colnames(cossim.df)[1] <- "ID"

## Remove high similarities of entries similar to itself
for(i in 1:nrow(cossim.df)) {
  cossim.df[i, i+1] = 0
}

## Find k most similar entries and add values from matrix to output df
for(col in 2:ncol(cossim.df)) {
  i = 0
  kmost <- sort(cossim.df[,col], index.return=TRUE, decreasing = TRUE)
  kmost <- lapply(kmost, `[`, kmost$x %in% head(kmost$x,k))
  for(coscol in 1:k) {
    cosine.df[col-1,coscol+i+1] <- kmost$ix[coscol]
    cosine.df[col-1,coscol+i+2] <- kmost$x[coscol]
    i = i + 1
  }
}
```

## Jaccard Similarity Function

```{r jaccard, include=FALSE}
## Convert data frame to matrix
newincome.m <- data.matrix(newincome.df)

## Create adjustable k
k <- 5 # Adjust k here

## Create a new data frame containing jaccard similarities and output values
jaccard.m <- matrix(nrow=(nrow(newincome.df)+1), ncol=(nrow(newincome.df)+1))
jaccard.m[2:nrow(jaccard.m),1] <- newincome.df$ID
jaccard.m[1,2:nrow(jaccard.m)] <- newincome.df$ID
jaccard.df <- data.frame(income.df$ID)
jaccard.df[,2:((k*2)+1)] <- 0

## Rename columns
names(jaccard.df)[seq(2,ncol(jaccard.df),2)] <- "Index"
names(jaccard.df)[seq(1,ncol(jaccard.df),2)] <- "Proximity"
names(jaccard.df)[1] <- "ID"

## Create a function for calculating euclidean similarity of two distinct data entries
jaccard <- function(indi, indj) {
  bmask <- rep(1, 41)
  xVec <- as.numeric(newincome.df[indi,2:42])
  yVec <- as.numeric(newincome.df[indj,2:42])
  M01 <- sum((!xVec&bmask)&(yVec&bmask))
  M10 <- sum((xVec&bmask)&(!yVec&bmask))
  M00 <- sum(!xVec & !yVec)
  M11 <- sum(xVec & yVec)
  simJ <- (M11)/(M01+M10+M11)
  return(simJ)
}

## Perform function for each entry
for (rowi in 1:nrow(newincome.df)) {
  for (rowj in 1:nrow(newincome.df)) {
    jaccard.m[rowi+1, rowj+1] <- jaccard(rowi, rowj)
  }
}

## Convert matrix to dataframe
jacsim.df <- data.frame(jaccard.m)
jacsim.df[1,1] = 0
colnames(jacsim.df) <- jacsim.df[1,]
jacsim.df <- jacsim.df[-1,]
colnames(jacsim.df)[1] <- "ID"

## Remove high similarities of entries similar to itself
for(i in 1:nrow(jacsim.df)) {
  jacsim.df[i, i+1] = 0
}

## Find k most similar entries and add values from matrix to output df
for(col in 2:ncol(jacsim.df)) {
  i = 0
  kmost <- sort(jacsim.df[,col], index.return=TRUE, decreasing = TRUE)
  kmost <- lapply(kmost, `[`, kmost$x %in% head(kmost$x,k))
  for(coscol in 1:k) {
    jaccard.df[col-1,coscol+i+1] <- kmost$ix[coscol]
    jaccard.df[col-1,coscol+i+2] <- kmost$x[coscol]
    i = i + 1
  }
}
```

The distribution of proximities between each example and its first nearest neighbor follows a similar shape. However, the data will be distributed among different proximity values, since two different proximity functions were used in the program. When k=5, the application will be gathering proximity values from the tail of the distribution containing the total proximity values for each entry. As “k” increases, the application will be gathering proximity values from the area closest to the mean. Essentially, the total proximity values for each entry follows a normal distribution, and the distribution of the gathered proximity values will begin to follow this normal distribution as “k” increases. Additionally, the class attribute seems to follow a similar pattern, although it was not used in the proximity function.

After running the program, we can see that a few data entries are very similar to each other, according to both the cosine and jaccard proximity functions. Some data entries are perfectly similar to other data entries, since continuous variables were manipulated to represent variables of categorical type after data preprocessing. For example, the entry with ID=9364 is perfectly similar to the entry with ID=12510, according to the Jaccard proximity function (Figure 7). Although a few of the variables from the original dataset are not exactly equal, the overall proximity of these values prove the effectiveness of the approach taken to transform variables and group certain values together during data preprocessing. 

## Export Data Frames
```{r export}
## Export cosine similarity output
write.csv(cosine.df, file = "CosineSimilarityOutput.csv")

## Export jaccard similarity output
write.csv(jaccard.df, file = "JaccardSimilarityOutput.csv")
```

